\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{multicol}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue
    }
%\usepackage[letterpaper, total={6in, 8in}]{geometry}
\usepackage[letterpaper]{geometry}
\geometry{verbose, tmargin=0.8in, bmargin=0.8in, lmargin=1in, rmargin=1in}
%\title{CIS545: Term Project Proposal}
%\author{Vijay Yevatkar, Saumya Suvarna}
%\date{November 2021}
\begin{document}
\begingroup  
  \centering
  \LARGE CIS545: Term Project Proposal\\[1.5em]
  \large Vijay Yevatkar, Saumya Suvarna\par
\endgroup
%\maketitle
\section{Data Source}
\href{http://www.murderdata.org/}{Murder Accountability Project}\\
“More than 256,000 Americans have perished in unsolved homicides committed since 1980 - more than the combined death toll of all U.S. military actions since World War II.”
\begin{itemize}
    \item With no law enforcement agency in America assigned to monitor failed homicide investigations, this project was set up to track unsolved homicides nationwide. 
    \item It is dedicated to educating Americans on the importance of accurately accounting for unsolved homicides within the United States. 
    \item The project provides CSV files summarizing all homicides and homicide clearances reported from year 1965 to 2020.
    \item The size of the data is a little shy of a million records - 827220
\end{itemize}

\section{Project Plan}
Our project will be roughly divided into the following sections:
\begin{enumerate}
    \item Data Wrangling\\
    This section will involve:
        \begin{itemize}
            \item Cleaning the data.
            \item Data preprocessing.
            \item Scaling, normalization, encoding values.
            \item \textbf{(Stretch Goal)} Merging data from various sources for additional insights.
        \end{itemize}
    \item Exploratory Data Analytics\\
    This section will involve:
        \begin{itemize}
            \item Visual representations of interesting aspects of the data.
            \item General statistics to get a better understanding of the overall data.
        \end{itemize}
    \item Modelling and fine tuning\\
    This section will involve:
        \begin{itemize}
            \item Selection of a model.
            \item Evaluating the performance.
            \item \textbf{(Stretch Goal)} Comparison between models that might give equivalent or better performance based on the data.
        \end{itemize}
    \item Bias Detection and Mitigation\\
    This section will involve:
        \begin{itemize}
            \item Evaluating if the model is making biased decisions (for the purpose of this project we will only consider based on race/gender).
            \item Mitigating these biases.
            \item \textbf{(Stretch Goal)} Debiasing the dataset.\\\\
        \end{itemize}
    \item Analysis\\
    This section will involve:
        \begin{itemize}
            \item Performance analysis of the model based on hyperparameter tuning.
            \item Performance comparison between the unbiased and biased models and the bias performance tradeoff.
        \end{itemize}
    \item \textbf{(Stretch Goal)} Interpretability\\
    This section will involve evaluating the interpretability of the model and a detailed analysis of performance vs interpretability tradeoff.
\end{enumerate}

\section{Objectives}
\begin{itemize}
    \item Our main objective of this project is to study the propagation of bias through a machine learning pipeline and the different ways we can mitigate it at various levels.
    \item At the end of this project we want to have made a prediction on a dataset that contains sensitive information, ensure that it is not biased and perform an analysis on the tradeoff of accuracy vs bias in the model.
    \item Based on the structure of our dataset, ensemble models will be our first choice of models to work with.
\end{itemize}

\section{Group Member Responsibilities}
\begin{multicols}{2}
\begin{itemize}
    \item Saumya Suvarna
    \begin{itemize}
        \item Exploratory Data Analytics.
        \item Bias detection and mitigation.
        \item \textbf{(Stretch Goal)}Interpretability
    \end{itemize}
    \item Vijay Yevatkar
    \begin{itemize}
        \item Data Preprocessing.
        \item Modelling and fine tuning.
        \item Analysis.
    \end{itemize}
\end{itemize}
\end{multicols}

\section{Problem Interest}
The problem is interesting for the following reasons:
\begin{itemize}
    \item We are working on real data collected.
    \item Many AI applications make biased predictions based on sex, race etc and we want to explore the complexity of ensuring that there is a reasonable bias vs accuracy tradeoff.
    \item Mitigate choices that are motivated by bias and ensure a degree of explainability for our model. 
    \item Ensure these steps can be easily replicated for a wider variety of applications to ensure no significantly biased predictions in any space.
\end{itemize}
\section{Potential Challenges}
\begin{itemize}
    \item One of the major challenges we anticipate with this project is a significant drop in performance with the debiased data/model. 
    \item We might also face the lack of resources in this particular domain if we run into issues.
    \item Would we be inducing an unknown bias by removing a known bias? 
\end{itemize}
\section{TA}
We would like to be assigned to Sally Hu as our TA for this project.

\end{document}
